\documentclass[../Main.tex]{subfiles}

\begin{document}
\chapter*{Preface}

\quad Physics has often been taught in a disconnected form. What begins with introductory classes in high school evolves into a mess of courses and textbooks during undergraduate education. Each subject is presented separately with its own assumptions, notation, and philosophy. The natural world, however, does not operate in fragments. The same underlying structures govern every phenomenon.

This book was written with one goal in mind: to bridge the fragmented teaching of physics. It aims to present physics, at an undergraduate level, as a single and coherent framework. Everything in the book is derived from first principles and strong mathematical rigor.

As a result, the book is a bit of a misnomer, it is not meant for all undergraduate students. It is meant for the student who wants to understand, solve, and live physics. At a minimum, the book expects one to be proficient in analysis and algebra, but more importantly, curiosity and patience. Physics rewards those who learn to sit with difficulty, to follow derivations to their logical end, and to appreciate the unity beneath the diversity of its equations.

My hope is two-fold. My first hope is that this book reaches its intended audience: students who intend to understand the world with more rigor than others. My second hope is that this book is enjoyed by its intended audience: as a tool to sharpen their physics skill.

Alright, I think I've scared, you enough now! On a personal note, I've hated waiting for the physics concepts to "get interesting." I've taken care to introduce physics in a way that shows its beauty quite quickly and evidently. I hope you finish the book with the same level of love of physics that I have.


\end{document}